Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline

### 2014-02-10 paper.tex
Updated reviewer comments

This displacement mechanism is capable of transferring charge between electrical and biological systems without damaging electrodes or tissue. \cite{Horch2004}
This behaviour is modelled by way of a Constant Phase Element (CPE); also known as a fractional capacitor.
}
For simulation purposes this element is realised as an array of R-C branches.~\cite{ScottSingle2013,Morrison59,Elwakil10} We measure displacement currents by sweeping the frequency of a sinusoidal stimulus across electrodes 2 and 7 of the Octrode while measuring the voltage between electrodes 2 and 3 (numbering shown in Fig.~\ref{fig:octrode}). 
This allows for the measurement of a single interface impedance in series with the resistance presented by the bulk solution. 
\begin{figure}
    \begin{center}
        \includegraphics{graphics/chargeInjectionVsFrequency.pdf}
    \end{center}
    \caption{\color{blue} Charge injected during CPE measurment for each of the six concentrations of PBS. Calculated using an effective surface area of 14\thinspace mm$^{2}$}
    \label{fig:chargeInjectionVsFrequency}
\end{figure}
{
    \color{blue}
Current density during measurements peaked at $11.5\thinspace A/m^{2}$ at 10\thinspace kHz in 1.0X PBS and fell as low as $0.0141\thinspace A/m^{2}$ at 50\thinspace mHz in 0.025X PBS. The stimulus waveform was set at 300\thinspace mV (peak) at each measurement point. The measurement insturments used are the same as those used to measure the inter-electrode resistance.
}
{
    \color{blue)
    % Item 6
    Figure~\ref{fig:CPE_Magnitude} shows the magnitude of the measured response of each of the six PBS concentrations accompanied by simulation results. Figure~\ref{fig:CPE_Phase} shows the phase response for the same measurements, again with the simulated response at each concentration. The simulated responses are generated using the optimised parameter values presented in Table~\ref{tab:CPEparams}. Those values were chosen as they give the closest match to the measured data. 
}
As shown in the interface model schematic (figure~\ref{fig:schematic}), the interface contains its own internal series resistance ($R_{S}$). The resistance seen in series with the CPE will therefore be the sum of both the resistance due to the bulk resistivity of the fluid and $R_{S}$, to which we will refer as $R_{S+bulk}$.
In figure~\ref{fig:CPE_Magnitude}, the slope and magnitude of the CPE are visible below 1\thinspace Hz whereas $R_{S+bulk}$  dominates above 1\thinspace kHz; it is clear that the two do not scale similarly.
As the impedance of the CPE and $R_{S+bulk}$ are separable with frequency, these measurements can be used to determine the CPE parameters and $R_{S}$.

@@

In \cite{Cogan2008} it has been noted that results obtained using cyclic voltammetry are dependant on, among other factors, the measurement sweep rate. This indicates a lack of isolation between displacement and Faradaic mechanisms. Our measurements were repeated using wait-times of 4, 16, 32 and 64 seconds between voltage increments, allowing us to determine the effect of sweep rate. From these measurements and those of a separate investigation\footnote{We took measurements of the interface's response to voltage steps when left to settle for 10 000 seconds, of which the first 250\thinspace seconds are shown in figure~\ref{fig:CPE_currentVsTime}. Those measurements show a trend of decreasing current stabalisation time as the overpotential steps increase. Sixty four seconds after stepping from 0.64\thinspace V to 0.72\thinspace V, the current has stabalised to $95\%$ of its value at 10\,000 seconds. At overpotential steps for higher voltages, the settling time is further reduced. These measurements were made in still solution and we expect stirring will further reduce settling time.} we concluded that a wait-time of 64 seconds was sufficient.
\begin{figure}
    \begin{center}
        \includegraphics{graphics/CPE_currentVsTime}
    \end{center}
    \caption{Current versus time after a stepped voltage increment across a pair of interfaces. Time before 64 seconds is highlighted in grey. Measurements were taken with a wait time of 10,000 seconds between increments.}
    \label{fig:CPE_currentVsTime}
\end{figure}

@@

%\begin{figure}
%    \begin{center}
%        \includegraphics{graphics/faradaic_currentVsVoltageIEEE}
%    \end{center}
%    \caption{Simulated response of the interface model compared to measurement from the electrodes with a stepped voltage overpotential. Measurement points show current values taken at the $64^{th}$ second after each voltage increment.}
%    \label{fig:faradaic_currentVsVoltage}
%\end{figure}

@@

{
    \color{blue}
Figure~\ref{fig:faradaic_currentVsTime} shows a subset of the data used in Fig.~\ref{fig:StepResponse_Faradaic} (the 1.0X PBS trace) to show the fit between simulated and measured data. Here the simulation contains fitted diode parameters, shown in table~\ref{tab:FaradaicParams}, which give rise to the exponentially increasing steady state current.
It is apparent from this trace that the temporal response of the CPE does not match the transient decay curves seen in the Faradaic measurements, something the CPE should predict. These transient decay curves appear to have time-constant that decreases with each subsequent measurement. This variable time-constant phenomena is especially clear when looking back at figure~\ref{fig:CPE_currentVsTime}. Although the Faradaic response transients appear to approach the time-constant predicted by the CPE, there is clearly behaviour not completely described by the CPE.

We hope to extend this model into a diffusion controlled conduction scenario and extend the CPE element to mimic the variable time-constant behaviour in a follow up paper.
}
\section{In-vivo measurements of sheep}

@@

page 2, line 52. "the magnitude shifts relatively little". However, In Fig. 4, logaritmic scale is used, and the change from 1.0X to 0.025X is almost 50k at the lowest frequencies. 
please clarify this point.\\
{\color{OliveGreen}
    Item 12\\
    Have reworded this slightly to convey that the shift in the CPE response is small relative to the shift in the resistance in the solution bulk. We are referring to a relative difference, a factor.
    {\color{red} I don't know how better to explain this, The difference between 9 000 000 and 1 000 000 000 is small compared to the difference between 20 and 2000}
}\\

page 3, in text/ table III. The parameters (m, k..) and their meanings should be described more comprehensively, as well.\\
{\color{OliveGreen} 
    Item 13\\
    {\color{Red} Describe the parameters m and k in the resistor network (pull from Jonathans paper and sub-references)}
    No response yet.
}\\

@@

page 4. Referencing (in the text) to the Fig. 10 is lacking (?).\\
{\color{OliveGreen}
    Item 15\\
    {\color{Red} Include a reference to Figure 10 in the document}
    No response yet.
}\\

Finally, and most importantly, the part of in vivo measurements are very limited since only one spinal channel measurement has been performed. My opinion is that this manuscript could be even better without this uncertain (n=1) result and thus I recommend to leave this part out.\\
{\color{OliveGreen}
    Item 16\\ 
    {\color{Red} Leave sheep in but perhaps show the second measurement (day 1) if possible (4th reviewer found sheep correlation to PBS result interesting). Would a separate sheep measurement graph be of use showing all sheep results? Or do we hold onto this for the better phantom paper?}
    No response yet.
}\\
}

@@

1. P1, top RH col. ``0.1X'' ought to be explained when you first use this form.\\
{\color{OliveGreen}
    Item 20\\
    Yes, thank you, see amendment.
}\\

2. Should the resistor model of the volume conductor from [4] be shown in this paper? It would make the paper easier to understand for your reader if you showed the model. If you are not going to do so, you might help them by saying something like ``Readers are advised to see [4] to understand the meaning of the parameters in Table II''.\\
{\color{OliveGreen}
    Item 21\\
    {\color{Red} Leave until the end, if we have the space then we put it in, otherwise put a better reference to ScottSingle2013}
    No response yet.
}\\

3. You say (p2 line 43) that there are 2 independent scaling factors. Is Rli independent or not? If not, why input it in Table II?\\
{\color{OliveGreen}
    Item 22\\
    {\color{Red} Mention that $R_{li}$ is not independent because of the geometric simplification of the resistor network. Mention that for more detail we suggest viewers read the relevant section of ScottSingle2013}
    No response yet.
}\\

@@

Technically, I found one issue with this paper. In Figure 11, the x-axis labels should have 2 digits after the decimal. As presently shown, each of the labels 0.6, 0.7, ..., 1.0 are repeated twice on the axis. Presumably, the second version of the labels should be 0.65, 0.75, ..., 1.05.\\
{\color{OliveGreen}
    Item 37\\
    {\color{Red} Definitely, how did I miss that?}
    No response yet.
}\\
}
\end{document}

>>>>

This displacement mechanism is capable of transferring charge between electrical and biological systems without damaging electrodes or tissue. \cite{Horch2004}
This behaviour is modelled by way of a Constant Phase Element (CPE); also known as a fractional capacitor.
}
{
    \color{blue}
For simulation purposes this element is realised as an array of R-C branches.~\cite{ScottSingle2013,Morrison59,Elwakil10} where the value of each R-C pair is chosen to set the slope of impedance relative to frequency. The placement of R-C pairs within the CPE is determined by the parameters $m$ and $k$, where $k$ determines the spacing, or density, of R-C branches and $m$ determines the resulting slope.
}
We measure displacement currents by sweeping the frequency of a sinusoidal stimulus across electrodes 2 and 7 of the Octrode while measuring the voltage between electrodes 2 and 3 (numbering shown in Fig.~\ref{fig:octrode}). 
This allows for the measurement of a single interface impedance in series with the resistance presented by the bulk solution. 
\begin{figure}
    \begin{center}
        \includegraphics{graphics/chargeInjectionVsFrequency.pdf}
    \end{center}
    \caption{\color{blue} Charge injected during CPE measurement for each of the six concentrations of PBS. Calculated using an effective surface area of 14\thinspace mm$^{2}$}
    \label{fig:chargeInjectionVsFrequency}
\end{figure}
{
    \color{blue}
Current density during measurements peaked at $11.5\thinspace A/m^{2}$ at 10\thinspace kHz in 1.0X PBS and fell as low as $0.0141\thinspace A/m^{2}$ at 50\thinspace mHz in 0.025X PBS. The stimulus waveform was set at 300\thinspace mV (peak) at each measurement point. The measurement instruments used are the same as those used to measure the inter-electrode resistance.
}
{
    \color{blue)
        {\color{red} something weird going on here} Figure~\ref{fig:CPE_Magnitude} shows the magnitude of the measured response of each of the six PBS concentrations accompanied by simulation results. Figure~\ref{fig:CPE_Phase} shows the phase response for the same measurements, again with the simulated response at each concentration. The simulated responses are generated using the optimised parameter values presented in Table~\ref{tab:CPEparams}. Those values were chosen as they give the closest match to the measured data. 
}
As shown in the interface model schematic (figure~\ref{fig:schematic}), the interface contains its own internal series resistance ($R_{S}$). The resistance seen in series with the CPE will therefore be the sum of both the resistance due to the bulk resistivity of the fluid and $R_{S}$, to which we will refer as $R_{S+bulk}$.
In figure~\ref{fig:CPE_Magnitude}, the slope and magnitude of the CPE are visible below 1\thinspace Hz whereas $R_{S+bulk}$  dominates above 1\thinspace kHz; it is clear that the two do not scale similarly.
As the impedance of the CPE and $R_{S+bulk}$ are separable with frequency, these measurements can be used to determine the CPE parameters and $R_{S}$.

@@

In \cite{Cogan2008} it has been noted that results obtained using cyclic voltammetry are dependant on, among other factors, the measurement sweep rate. This indicates a lack of isolation between displacement and Faradaic mechanisms. Our measurements were repeated using wait-times of 4, 16, 32 and 64 seconds between voltage increments, allowing us to determine the effect of sweep rate. From these measurements and those of a separate investigation\footnote{We took measurements of the interface's response to voltage steps when left to settle for 10 000 seconds, of which the first 250\thinspace seconds are shown in figure~\ref{fig:CPE_currentVsTime}. Those measurements show a trend of decreasing current stabilisation time as the overpotential steps increase. Sixty four seconds after stepping from 0.64\thinspace V to 0.72\thinspace V, the current has stabalised to $95\%$ of its value at 10\,000 seconds. At overpotential steps for higher voltages, the settling time is further reduced. These measurements were made in still solution and we expect stirring will further reduce settling time.} we concluded that a wait-time of 64 seconds was sufficient.
\begin{figure}
    \begin{center}
        \includegraphics{graphics/CPE_currentVsTime}
    \end{center}
    \caption{Current versus time after a stepped voltage increment across a pair of interfaces. Time before 64 seconds is highlighted in grey. Measurements were taken with a wait time of 10,000 seconds between increments.}
    \label{fig:CPE_currentVsTime}
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics{graphics/faradaic_currentVsVoltageIEEE}
    \end{center}
    \caption{Simulated response of the interface model compared to measurement from the electrodes with a stepped voltage overpotential. Measurement points show current values taken at the $64^{th}$ second after each voltage increment.}
    \label{fig:faradaic_currentVsVoltage}
\end{figure}

@@

{
    \color{blue}
    Figure~\ref{fig:faradaic_currentVsTime} shows a subset of the data used in Fig.~\ref{fig:StepResponse_Faradaic} (the 1.0X PBS trace) to show the fit between simulated and measured data in the time domain. Here the simulation contains fitted diode parameters, shown in table~\ref{tab:FaradaicParams}, which give rise to the exponentially increasing steady state current.     
It is apparent from this trace that the temporal response of the CPE does not match the transient decay curves seen in the Faradaic measurements, something the CPE should predict. These transient decay curves appear to have time-constant that decreases with each subsequent measurement. This variable time-constant phenomena is especially clear when looking back at figure~\ref{fig:CPE_currentVsTime}. Although the Faradaic response transients appear to approach the time-constant predicted by the CPE, there is still behaviour not completely described by the CPE. 
The match between fitted diode parameters ($n$ and $i_{0}$) and the measurement data 64 seconds after each voltage increment has been plotted in figure~\ref{fig:faradaic_currentVsVoltage}. This graph plots current conducted at 64 seconds after each step in terms of voltage applied across electrodes 3 and 7.

In future we hope to extend this model into a diffusion controlled conduction scenario and extend the CPE element to better reproduce the time domain response, which we hope to publish in a follow-up paper.
}
\section{In-vivo measurements of sheep}

@@

page 2, line 52. "the magnitude shifts relatively little". However, In Fig. 4, logaritmic scale is used, and the change from 1.0X to 0.025X is almost 50k at the lowest frequencies. 
please clarify this point.\\
{\color{OliveGreen}
    Item 12\\
    Have reworded this slightly to convey that the shift in the CPE response is small relative to the shift in the resistance in the solution bulk. We are referring to a relative difference, a gain in conductivity as a factor of the current conductivity.
    {\color{red} Help! I don't know how better to explain this, The difference between 9 000 000 and 1 000 000 000 is small compared to the difference between 20 and 2000}
}\\

page 3, in text/ table III. The parameters (m, k..) and their meanings should be described more comprehensively, as well.\\
{\color{OliveGreen} 
    Item 13\\
    Sure, we have added a quick interpretation for the parameters $m$ and $k$ in the text.
}\\

@@

page 4. Referencing (in the text) to the Fig. 10 is lacking (?).\\
{\color{OliveGreen}
    Item 15\\
    Thank you, we have updated discussion of Faradaic Parameters to include the results from what is now figure~\ref{fig:faradaic_currentVsTime}.
}\\

Finally, and most importantly, the part of in vivo measurements are very limited since only one spinal channel measurement has been performed. My opinion is that this manuscript could be even better without this uncertain (n=1) result and thus I recommend to leave this part out.\\
{\color{OliveGreen}
    Item 16\\ 
    {\color{Red} Leave sheep in but perhaps show the second measurement (day 1) if possible (4th reviewer found sheep correlation to PBS result interesting). Would a separate sheep measurement graph be of use showing all sheep results? Jonathans response: Put more sheep data in, put some error bars on.}
    No response yet.
}\\
}

@@

1. P1, top RH col. ``0.1X'' ought to be explained when you first use this form.\\
{\color{OliveGreen}
    Item 20\\
    Fixed.
}\\

2. Should the resistor model of the volume conductor from [4] be shown in this paper? It would make the paper easier to understand for your reader if you showed the model. If you are not going to do so, you might help them by saying something like ``Readers are advised to see [4] to understand the meaning of the parameters in Table II''.\\
{\color{OliveGreen}
    Item 21\\
    {\color{Red} Leave until the end, if we have the space then we put it in, otherwise put a better reference to ScottSingle2013}
    We have advised the reader to read ScottSingle2013 to understand the interpretation of the mesh generation parameters, where they would find the radial volume model.
}\\

3. You say (p2 line 43) that there are 2 independent scaling factors. Is Rli independent or not? If not, why input it in Table II?\\
{\color{OliveGreen}
    Item 22\\
    Footnote added to sentence in question that describes our defining of $R_{sri}$ in terms of $R_{eri}$. 
}\\

@@

Technically, I found one issue with this paper. In Figure 11, the x-axis labels should have 2 digits after the decimal. As presently shown, each of the labels 0.6, 0.7, ..., 1.0 are repeated twice on the axis. Presumably, the second version of the labels should be 0.65, 0.75, ..., 1.05.\\
{\color{OliveGreen}
    Item 37\\
    Thank you very much. See updated figure (now figure~\ref{fig:faradaic_currentVsVoltage}).
}\\
}
\end{document}
